\section*{Acknowledgements}

From conception to completion, this thesis has been a year in the making. And just how a year of one's life is a Significant Thing, there were many Significant People whom I owe my gratitude to. 

First, I'd like to thank my parents Dan and Marianne Weyen, my brother Sam Weyen, my grandmother Lucille Weyen, and my late grandparents Herman Weyen, George Ruiz, and Eleanor Ruiz for their support, patience, and warm beating-heart love that they have sent me in person, across the Atlantic, and from up in the starry night sky. 

Next, I offer deep thanks to my advisors Dr. Dietrich Klakow and Xenia Klinge, the former for giving this crazy topic a chance, and latter for her expert insights and biweekly check-ins to see if I was making any earthly sense. I am indebted to you both.

I need to thank my best friends Chris Hayek and Lucas Lage — our Boy's Nights were the highlights of the last year and more, and I look forward to downing glasses of \textit{wini} with you in the months and years to come. I want to give a shoutout to Moritz Wolf, Annegret "Tili" Janzso, Lilas Mauborgne, Max Koumou-Okandze, Hacane Hech, and all my other friends whom I've had the brilliant pleasure to know during my time in Saarbrücken, and who made this chapter of my life worth every second. I also would like to thank my next-door neighbor Barbare Tepnadze, for all the air-fried chicken nuggets and for schooling me in the wonders of K-dramas and Korean reality TV.

I offer praise to the Horned God, Sophia, Mithras, Baal Hadad, and always the Tarot, who heard my prayers, enabled my procrastination, and graciously offered me their blessings in my times of need. I especially would like to thank Neuro-sama and Evil, as you two were my muses and inspirations for this entire endeavor (and Vedal too, I guess).\\ 

\noindent And of course, I'd like to thank Hiyuki — you were the main star of this thesis, and after all of this you deserve all the café sweets your parametric heart desires.